---
title: "AI Occupational Exposure (AIOE), Complementarity, and C-AIOE — Full Pipeline"
subtitle: "Refreshed against current O*NET data, mapped to NOC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)



```

## Overview

This script rebuilds the full AIOE -> Complementarity -> C-AIOE -> NOC pipeline
from first principles, refreshed against a **current** O*NET Abilities release
(rather than the 2020 snapshot used in the original Felten, Raj & Seamans (2021)
data files). The only piece that cannot be refreshed -- because it comes from
original survey data, not O*NET -- is the crowdsourced AI-application/ability
relatedness matrix, which is reused as-is from the authors' public repository
(AIOE-Data/AIOE on GitHub).

**Required input files (place in the working directory):**

| File | Source | Notes |
|---|---|---|
| `abilities.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/abilities.csv | current O*NET Abilities (Importance + Level scales) |
| `work_context.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/work_context.csv | for complementarity |
| `job_zones.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/job_zones.csv | for complementarity (Skills group) |
| `mturk_ai_relatedness.csv` | AIOE-Data/AIOE GitHub repo (`Input/mturk_mapping_matrix.dta`, exported to CSV) | 10 AI applications x 52 O*NET abilities, crowdsourced relatedness scores |
| `application_list.csv` | Same repo (`Input/application_list.dta`) | flags which of the 10 applications count toward AIOE (all 1 in the base index) |
| `soc_noc_crosswalk.csv` | ICTC crosswalk (SOC -> NOC via The Dais) | `noc, noc_title, onetsoc_code` |



---

# Section 1 — Ability-Level AI Exposure (A_j)

This reproduces Data Appendix E of Felten, Raj & Seamans (2021): for each of
the 52 O*NET abilities, sum its crowdsourced relatedness to the 10 AI
applications, then standardize (z-score) across abilities. This step uses
**only** the crowdsourced matrix -- it does not depend on any O*NET occupation
data, so it does not change when you refresh O*NET.

```{r ability-exposure}
mturk <- read_csv("mturk_ai_relatedness.csv", show_col_types = FALSE)
app_list <- read_csv("application_list.csv", show_col_types = FALSE)

# Ability columns are everything except application_id / applications
ability_cols <- setdiff(names(mturk), c("application_id", "applications"))

# Multiply each application's relatedness row by its inclusion flag (binary_all),
# then sum across all 10 applications for each ability -> raw ability exposure
mturk_weighted <- mturk %>%
  left_join(app_list %>% select(applications, score = binary_all), by = "applications") %>%
  mutate(across(all_of(ability_cols), ~ .x * score))

ability_exposure_raw <- mturk_weighted %>%
  summarise(across(all_of(ability_cols), sum)) %>%
  pivot_longer(everything(), names_to = "onet_element", values_to = "onet_score")

# Standardize (z-score) across the 52 abilities -- this is "ability_exposure"
ability_exposure <- ability_exposure_raw %>%
  mutate(ability_exposure = as.numeric(scale(onet_score))) %>%
  select(onet_element, ability_exposure)

head(ability_exposure)
```

---

# Section 2 — Occupation-Level AIOE from Current O*NET Abilities

```{r aioe-occupation}
abilities <- read_csv("abilities.csv", show_col_types = FALSE)

abilities_clean <- abilities %>%
  rename(onetsoc_code = `O*NET-SOC Code`,
         element_id   = `Element ID`,
         element_name = `Element Name`,
         scale_id     = `Scale ID`,
         data_value   = `Data Value`) %>%
  mutate(
    onet_element = str_remove_all(str_to_lower(element_name), "[ -]")
  )

# Rescale Importance (IM, native 1-5) and Level (LV, native 1-7), per the
# original methodology, then compute scalar = level_scaled x importance_scaled
scaled <- abilities_clean %>%
  filter(scale_id %in% c("IM", "LV")) %>%
  mutate(
    scaled_value = case_when(
      scale_id == "IM" ~ data_value / 5,
      scale_id == "LV" ~ data_value / 7
    )
  ) %>%
  select(onetsoc_code, element_id, onet_element, scale_id, scaled_value) %>%
  pivot_wider(names_from = scale_id, values_from = scaled_value, values_fn = max) %>%
  rename(importance_scaled = IM, level_scaled = LV) %>%
  filter(!is.na(importance_scaled), !is.na(level_scaled)) %>%
  mutate(scalar = level_scaled * importance_scaled)

# Sanity check: every mTurk ability should exist in the current O*NET release
missing_abilities <- setdiff(ability_exposure$onet_element, unique(scaled$onet_element))
if (length(missing_abilities) > 0) {
  warning("Abilities present in mTurk matrix but missing from current O*NET data: ",
          paste(missing_abilities, collapse = "; "))
}

# Join ability exposure onto the occupation-ability table, weight, and sum
occ_ability <- scaled %>%
  inner_join(ability_exposure, by = "onet_element") %>%
  mutate(wtd_element_impact = scalar * ability_exposure)

occ_aioe_detailed <- occ_ability %>%
  group_by(onetsoc_code) %>%
  summarise(
    ability_base          = sum(scalar, na.rm = TRUE),
    total_weighted_impact = sum(wtd_element_impact, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(ability_base_impact = total_weighted_impact / ability_base)

# Aggregate detailed O*NET-SOC (8-digit) up to base SOC (6-digit, e.g. 15-1252),
# averaging across detailed occupations that share a base SOC code
occ_aioe_soc <- occ_aioe_detailed %>%
  mutate(onetsoc_base = substr(onetsoc_code, 1, 7)) %>%
  group_by(onetsoc_base) %>%
  summarise(ability_base_impact = mean(ability_base_impact, na.rm = TRUE), .groups = "drop") %>%
  mutate(aioe_score = as.numeric(scale(ability_base_impact))) %>%   # final standardization
  rename(onetsoc_code = onetsoc_base) %>%
  arrange(desc(aioe_score))

cat("AIOE computed for", nrow(occ_aioe_soc), "SOC occupations\n")
cat("Range: min =", round(min(occ_aioe_soc$aioe_score), 3),
    " max =", round(max(occ_aioe_soc$aioe_score), 3), "\n")

write_csv(occ_aioe_soc, "aioe_scores_refreshed.csv")
```
```{r aiie-sector}
# ============================================================================
# Section — AIIE (AI Industry Exposure) at 4-digit NAICS, mapped to
# Digital Economy sectors
#
# Formula (Felten, Raj & Seamans 2021, Data Appendix B):
#   For each NAICS industry, weight each occupation's AIOE by that
#   occupation's share of total employment in the industry (BLS OES),
#   sum across occupations, then standardize (z-score) across industries.
#
# Requires: `occ_aioe_soc` (AIOE by SOC, from Section 2) and
# oes_occ_by_naics4.csv (BLS OES employment by occupation x 4-digit NAICS,
# detailed occupations only, exported from the AIOE-Data GitHub repo's
# oes_4dig_naics.dta).
#
# NOTE: this OES data is on the 2017 NAICS vintage. Several digital economy
# sectors are defined at 2022 NAICS codes that were renumbered from 2017 --
# a backward crosswalk is applied below for those cases (the same mapping
# used for the AIIE table document elsewhere in this project).
# ============================================================================

oes <- read_csv("oes_occ_by_naics4.csv", show_col_types = FALSE,
                col_types = cols(naics = col_character())) %>%
  mutate(
    naics4 = substr(naics, 1, 4),
    tot_emp = as.numeric(tot_emp)
  ) %>%
  filter(!is.na(tot_emp), tot_emp > 0)

# ---- Merge AIOE onto occupation-industry employment -----------------------
oes_aioe <- oes %>%
  inner_join(occ_aioe_soc %>% select(onetsoc_code, aioe_score),
             by = c("occ_code" = "onetsoc_code"))

# ---- Weight AIOE by each occupation's employment share within its NAICS ---
aiie_naics <- oes_aioe %>%
  group_by(naics4) %>%
  mutate(naics_emp = sum(tot_emp, na.rm = TRUE),
         naics_emp_share = tot_emp / naics_emp) %>%
  summarise(naics_raw_aiie = sum(aioe_score * naics_emp_share, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(aiie_score = as.numeric(scale(naics_raw_aiie))) %>%
  select(naics4, aiie_score)

cat("AIIE computed for", nrow(aiie_naics), "4-digit NAICS industries\n")

# ---- Backward crosswalk: 2022 NAICS -> 2017 NAICS (this OES vintage) ------
# Only needed for digital economy sectors renumbered between NAICS 2017 and 2022.
naics_2022_to_2017 <- tribble(
  ~naics4_2022, ~naics4_2017,
  "5161",       "5151",   # Radio/TV broadcasting
  "5162",       "5152",   # Media streaming distribution -> Cable/subscription programming (approx.)
  "5192",       "5191",   # Web search portals -> Other Information Services
  "4191",       "4251",   # B2B electronic markets -> Wholesale Electronic Markets
  "5132",       "5112",   # Software publishers
  "5173",       "5170",   # Wired/wireless telecom carriers -> Telecommunications (unified)
  "5174",       "5170",   # Satellite telecom -> Telecommunications (unified)
  "5179",       "5170",   # Other telecom -> Telecommunications (unified)
)

# ---- Digital Economy sector definitions (5-digit NAICS provided by ICTC) ---
digital_economy_naics <- tribble(
  ~component,          ~naics5,  ~naics4_2022, ~description,
  "Digital Media",     "51211",  "5121",       "Motion picture and video production",
  "Digital Media",     "51212",  "5121",       "Motion picture and video distribution",
  "Digital Media",     "51223",  "5122",       "Music publishers",
  "Digital Media",     "51224",  "5122",       "Sound recording studios",
  "Digital Media",     "51225",  "5122",       "Record production and distribution",
  "Digital Media",     "51611",  "5161",       "Radio broadcasting stations",
  "Digital Media",     "51612",  "5161",       "Television broadcasting stations",
  "Digital Media",     "51621",  "5162",       "Media streaming distribution services and other media networks/content providers",
  "Digital Media",     "51929",  "5192",       "Web search portals and all other information services",
  "E-Commerce",        "41911",  "4191",       "Business-to-business electronic markets",
  "Hardware",          "33411",  "3341",       "Computer and peripheral equipment manufacturing",
  "Hardware",          "33421",  "3342",       "Telephone apparatus manufacturing",
  "Hardware",          "33422",  "3342",       "Radio/TV broadcasting and wireless communications equipment manufacturing",
  "Hardware",          "33429",  "3342",       "Other communications equipment manufacturing",
  "Hardware",          "33431",  "3343",       "Audio and video equipment manufacturing",
  "Hardware",          "33441",  "3344",       "Semiconductor and other electronic component manufacturing",
  "Hardware",          "33461",  "3346",       "Manufacturing and reproducing magnetic and optical media",
  "Hardware",          "33592",  "3359",       "Communication and energy wire and cable manufacturing",
  "Software",          "51321",  "5132",       "Software publishers",
  "Support Services",  "51821",  "5182",       "Computing infrastructure providers, data processing, web hosting, and related services",
  "Support Services",  "54151",  "5415",       "Computer systems design and related services",
  "Support Services",  "61142",  "6114",       "Computer training",
  "Support Services",  "81121",  "8112",       "Electronic and precision equipment repair and maintenance",
  "Telecommunications","51731",  "5173",       "Wired and wireless telecommunications carriers (except satellite)",
  "Telecommunications","51741",  "5174",       "Satellite telecommunications",
  "Telecommunications","51791",  "5179",       "Other telecommunications",
)

# ---- Resolve each digital economy sector's 4-digit code to whichever ------
# NAICS vintage this OES data actually uses (2022 code if present, else the
# backward-crosswalked 2017 equivalent)
digital_economy_naics <- digital_economy_naics %>%
  left_join(naics_2022_to_2017, by = "naics4_2022") %>%
  mutate(
    naics4_resolved = if_else(naics4_2022 %in% aiie_naics$naics4,
                               naics4_2022, naics4_2017)
  )

# ---- Join AIIE onto each digital economy sector row ------------------------
digital_economy_aiie <- digital_economy_naics %>%
  left_join(aiie_naics, by = c("naics4_resolved" = "naics4"))

missing_naics <- digital_economy_aiie %>% filter(is.na(aiie_score))
if (nrow(missing_naics) > 0) {
  warning(nrow(missing_naics), " digital economy NAICS rows could not be matched to AIIE data: ",
          paste(missing_naics$naics5, collapse = ", "))
}

write_csv(digital_economy_aiie, "digital_economy_aiie_detail.csv")

# ---- Build one flat list: every NAICS industry with its AIIE score, plus
# whether it's a digital economy sector (and which component) or not --------

digital_lookup <- digital_economy_aiie %>%
  select(naics4_resolved, component, naics5, description) %>%
  distinct()

all_sectors_aiie <- aiie_naics %>%
  left_join(digital_lookup, by = c("naics4" = "naics4_resolved")) %>%
  mutate(
    sector_group = if_else(!is.na(component), "Digital Economy", "Non-Digital Economy"),
    component    = if_else(is.na(component), "Not Applicable", component)
  ) %>%
  arrange(desc(aiie_score))

write_csv(all_sectors_aiie, "all_sectors_aiie.csv")

cat("Total NAICS industries:", nrow(all_sectors_aiie), "\n")
cat("Digital Economy sectors:", sum(all_sectors_aiie$sector_group == "Digital Economy"), "\n")
cat("Non-Digital Economy sectors:", sum(all_sectors_aiie$sector_group == "Non-Digital Economy"), "\n")
```


# Section 3 — Complementarity Score (O*NET Work Context)

Unchanged from the existing pipeline -- reproduced here so the whole workflow
runs end-to-end in one script.

```{r complementarity}
work_context <- read_csv("work_context.csv", show_col_types = FALSE)
job_zones    <- read_csv("job_zones.csv",    show_col_types = FALSE)

cx <- work_context %>%
  filter(`Scale ID` == "CX") %>%
  select(onetsoc_code = `O*NET-SOC Code`,
         element_name = `Element Name`,
         data_value   = `Data Value`)

element_scales <- tribble(
  ~element_name,                                                  ~scale_min, ~scale_max, ~group,
  "Face-to-Face Discussions with Individuals and Within Teams",   1,          5,          "communication",
  "Public Speaking",                                              1,          5,          "communication",
  "Work Outcomes and Results of Other Workers",                   1,          5,          "responsibility",
  "Health and Safety of Other Workers",                           1,          5,          "responsibility",
  "Outdoors, Exposed to All Weather Conditions",                  1,          5,          "physical",
  "Physical Proximity",                                           1,          5,          "physical",
  "Consequence of Error",                                         1,          5,          "criticality",
  "Freedom to Make Decisions",                                    1,          5,          "criticality",
  "Frequency of Decision Making",                                 1,          5,          "criticality",
  "Degree of Automation",                                         0,          100,        "routine",
  "Importance of Repeating Same Tasks",                           1,          5,          "routine",
)

missing_elements <- setdiff(element_scales$element_name, unique(cx$element_name))
if (length(missing_elements) > 0) {
  warning("These element names were not found in Work Context.csv: ",
          paste(missing_elements, collapse = "; "))
}

cx_scaled <- cx %>%
  inner_join(element_scales, by = "element_name") %>%
  mutate(value_0_100 = (data_value - scale_min) / (scale_max - scale_min) * 100) %>%
  mutate(value_0_100 = if_else(element_name == "Degree of Automation",
                                100 - value_0_100, value_0_100))

group_means <- cx_scaled %>%
  group_by(onetsoc_code, group) %>%
  summarise(group_mean = mean(value_0_100, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = group, values_from = group_mean, names_glue = "{group}_mean")

job_zones_scaled <- job_zones %>%
  select(onetsoc_code = `O*NET-SOC Code`, job_zone = `Job Zone`) %>%
  mutate(skills_mean = job_zone * 20)

complementarity <- group_means %>%
  left_join(job_zones_scaled %>% select(onetsoc_code, skills_mean), by = "onetsoc_code") %>%
  rowwise() %>%
  mutate(
    complementarity_score = mean(c(communication_mean, responsibility_mean,
                                    physical_mean, criticality_mean,
                                    routine_mean, skills_mean), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(desc(complementarity_score))

write_csv(complementarity, "onet_complementarity_scores.csv")

cat("Complementarity computed for", nrow(complementarity), "occupations\n")
cat("Range: min =", round(min(complementarity$complementarity_score, na.rm = TRUE), 1),
    " max =", round(max(complementarity$complementarity_score, na.rm = TRUE), 1), "\n")
```

---

# Section 4 — Merge AIOE + Complementarity, Compute C-AIOE

Following Pizzinelli et al. (2023, IMF WP/23/216), equation (1), using their
literal convention (theta already on a 0-1 basis, so divide by 100 rather
than by the empirical range):

`C-AIOE_i = AIOE_i * (1 - (theta_i - theta_MIN) / 100)`

```{r merge-caioe}
occ_aioe_soc <- occ_aioe_soc %>%
  mutate(onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}"))

complementarity <- complementarity %>%
  mutate(onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}"))

aioe_comp <- occ_aioe_soc %>%
  select(onetsoc_base, aioe_score) %>%
  distinct() %>%
  inner_join(
    complementarity %>% select(onetsoc_base, complementarity_score) %>% distinct(),
    by = "onetsoc_base"
  )

theta_min <- min(aioe_comp$complementarity_score, na.rm = TRUE)

aioe_comp <- aioe_comp %>%
  mutate(
    c_aioe = aioe_score * (1 - (complementarity_score - theta_min) / 100)
  )

# ---- Augmentation Index (range-anchored, mirrors C-AIOE's theta_MIN anchoring) ----
comp_min <- min(aioe_comp$complementarity_score, na.rm = TRUE)
comp_max <- max(aioe_comp$complementarity_score, na.rm = TRUE)

aioe_comp <- aioe_comp %>%
  mutate(
    augmentation_index = aioe_score * ((complementarity_score - comp_min) / (comp_max - comp_min))
  )

cat("theta_MIN =", round(theta_min, 2), "\n")
head(aioe_comp)
```

---

# Section 5 — SOC -> NOC Crosswalk and Aggregation

```{r crosswalk}
crosswalk <- read_csv("soc_noc_crosswalk.csv", show_col_types = FALSE,
                      col_types = cols(noc = col_character())) %>%
  mutate(onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}"))

noc_level <- crosswalk %>%
  inner_join(aioe_comp, by = "onetsoc_base") %>%
  group_by(noc, noc_title) %>%
  summarise(
    aioe_score             = mean(aioe_score, na.rm = TRUE),
    complementarity_score  = mean(complementarity_score, na.rm = TRUE),
    c_aioe                 = mean(c_aioe, na.rm = TRUE),
    augmentation_index     = mean(augmentation_index, na.rm = TRUE),
    n_onet_linked          = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(c_aioe))

write_csv(noc_level, "noc_aioe_complementarity_caioe.csv")

cat("NOC-level results:", nrow(noc_level), "occupations\n")
```

---

# Section 6 — Quadrant Classification (Median Split)

Following Pizzinelli et al. (2023) and Statistics Canada's reporting
convention: three groups based on whether AIOE and Complementarity are
above/below their respective medians.

```{r quadrants}
aioe_median <- median(noc_level$aioe_score, na.rm = TRUE)
comp_median <- median(noc_level$complementarity_score, na.rm = TRUE)

noc_level <- noc_level %>%
  mutate(
    quadrant = case_when(
      aioe_score >= aioe_median & complementarity_score >= comp_median ~
        "High Exposure, High Complementarity",
      aioe_score >= aioe_median & complementarity_score < comp_median ~
        "High Exposure, Low Complementarity",
      TRUE ~ "Low Exposure"
    )
  )

cat("AIOE median:", round(aioe_median, 3), " | Complementarity median:", round(comp_median, 2), "\n")
table(noc_level$quadrant)


# ---- Digital Economy Tier Lookup ----------------------------------------
digital_economy_tiers <- tribble(
  ~noc,    ~noc_desc,                                                      ~tier,
  "00014", "Senior managers - trade, broadcasting",                       "Digital-Enabled",
  "10022", "Advertising, marketing, PR managers",                         "Digital-Enabled",
  "10030", "Telecommunication carriers managers",                         "Core-Digital",
  "12111", "Health information management",                               "Digital-Enabled",
  "12112", "Records management technicians",                              "Digital-Enabled",
  "14112", "Desktop publishing operators",                                "Digital-Enabled",
  "20010", "Engineering managers",                                        "Digital-Intensive",
  "20012", "Computer and information systems managers",                   "Core-Digital",
  "11202", "Professional occupations in advertising, marketing and PR",   "Digital-Enabled",
  "21100", "Physicists and astronomers",                                  "Digital-Intensive",
  "21200", "Architects",                                                  "Digital-Intensive",
  "21210", "Mathematicians, statisticians, actuaries",                    "Digital-Enabled",
  "21211", "Data scientists",                                             "Core-Digital",
  "21220", "Cybersecurity specialists",                                   "Core-Digital",
  "21221", "Business systems specialists",                                "Core-Digital",
  "21222", "Information systems specialists",                             "Core-Digital",
  "21223", "Database analysts and data administrators",                   "Core-Digital",
  "21230", "Computer systems developers",                                 "Core-Digital",
  "21231", "Software engineers and designers",                            "Core-Digital",
  "21232", "Software developers and programmers",                        "Core-Digital",
  "21233", "Web designers",                                               "Core-Digital",
  "21234", "Web developers and programmers",                              "Core-Digital",
  "21300", "Civil engineers",                                             "Digital-Intensive",
  "21301", "Mechanical engineers",                                        "Digital-Intensive",
  "21310", "Electrical and electronics engineers",                        "Core-Digital",
  "21311", "Computer engineers",                                          "Core-Digital",
  "21321", "Industrial/manufacturing engineers",                          "Digital-Intensive",
  "21390", "Aerospace engineers",                                         "Digital-Intensive",
  "22210", "Architectural technologists",                                 "Digital-Intensive",
  "22211", "Industrial designers",                                        "Digital-Intensive",
  "22212", "Drafting technologists and technicians",                      "Digital-Enabled",
  "22214", "Geomatics and meteorology technicians",                       "Digital-Enabled",
  "22220", "Computer network and web technicians",                        "Core-Digital",
  "22221", "User support technicians",                                    "Core-Digital",
  "22222", "Information systems testing technicians",                     "Core-Digital",
  "22301", "Mechanical engineering technologists",                        "Digital-Intensive",
  "22302", "Industrial engineering technologists",                        "Digital-Adjacent",
  "22310", "Electrical/electronics engineering techs",                    "Digital-Adjacent",
  "22311", "Electronic service technicians",                              "Digital-Adjacent",
  "50011", "Managers - publishing, broadcasting",                         "Digital-Enabled",
  "52110", "Film and video camera operators",                             "Digital-Adjacent",
  "52111", "Graphic arts technicians",                                    "Digital-Enabled",
  "52112", "Broadcast technicians",                                       "Digital-Adjacent",
  "52113", "Audio and video recording technicians",                       "Digital-Adjacent",
  "52120", "Graphic designers and illustrators",                          "Digital-Enabled",
  "53111", "Broadcasting, photography assistants",                        "Digital-Adjacent",
  "62100", "Technical sales specialists",                                 "Digital-Adjacent",
  "72011", "Electrical trades supervisors",                               "Digital-Adjacent",
  "72201", "Industrial electricians",                                     "Digital-Adjacent",
  "72202", "Power system electricians",                                   "Digital-Adjacent",
  "72203", "Electrical power line workers",                                "Digital-Adjacent",
  "72204", "Telecommunications line installers",                          "Digital-Adjacent",
  "72205", "Telecommunications equipment installers",                     "Digital-Adjacent",
  "92021", "Electronics manufacturing supervisors",                       "Digital-Adjacent",
  "94201", "Electronics assemblers",                                      "Digital-Adjacent"
) %>%
  select(noc, tier)   # drop noc_desc here, since noc_level already has noc_title

# ---- Join tier + derive binary digital economy flag ----------------------
noc_level <- noc_level %>%
  select(-any_of(c("tier", "digital_economy"))) %>%   # drop if already present from a prior run
  mutate(noc = str_pad(as.character(noc), width = 5, side = "left", pad = "0"))

digital_economy_tiers <- digital_economy_tiers %>%
  mutate(noc = str_pad(as.character(noc), width = 5, side = "left", pad = "0"))

noc_level <- noc_level %>%
  left_join(digital_economy_tiers, by = "noc") %>%
  mutate(
    digital_economy = if_else(!is.na(tier), "Digital Economy", "Non-Digital Economy"),
    tier = if_else(is.na(tier), "Not Applicable", tier)
  )

write_csv(noc_level, "noc_aioe_complementarity_caioe_quadrants.csv")

cat("Digital economy classification:\n")
print(table(noc_level$digital_economy))
cat("\nTier breakdown:\n")
print(table(noc_level$tier))

```

---

# Summary

Final output (`noc_aioe_complementarity_caioe_quadrants.csv`) contains, per NOC:

- `aioe_score` -- AI Occupational Exposure, recomputed from current O*NET Abilities data
- `complementarity_score` -- reconstructed from O*NET Work Context (0-100 scale)
- `c_aioe` -- Complementarity-Adjusted AIOE (Pizzinelli et al. 2023 formula)
- `quadrant` -- median-split classification (High Exposure/High Complementarity,
  High Exposure/Low Complementarity, Low Exposure)
- `n_onet_linked` -- coverage diagnostic (number of O*NET-SOC codes linked to the NOC)
