---
title: "OECD AI Capability Gap Index by Occupation"
subtitle: "Domain-level gaps, aggregate index, AIOE/Complementarity, NOC mapping -- single self-contained pipeline"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tibble)
library(purrr)
```

## Overview

This script builds the OECD "AI Capability Gap" index by occupation, following
OECD AI Papers No. 59 ("The OECD AI Exposure Measure"). For each occupation `j`
and each of 9 capability domains `i`:

```
CG_ji = max(0, H_ji - AI_i)
```

where `H_ji` is the occupation's required capability level (0-5 scale, O*NET-derived)
and `AI_i` is the **current** AI capability level in that domain (a single fixed
value per domain, supplied below -- not occupation-specific). The domain gaps are
combined into a total index:

```
CG_j = sum_i ( CG_ji * Importance_ji )
```

We do not have OECD's original per-occupation importance weights (their Table 6.2
factor-analysis crosswalk), so Section 4 builds two documented, defensible
alternatives -- an equal-weighted average and a set of weights fit by regression --
and lets us compare them before picking one.

Everything here (domain gaps, total index, reversed/normalized score) is
**recomputed from scratch** from the raw job-demand (H_ji) values and our own
AI_i assumptions. The "AI Capability Gap Index" and per-domain "Capability Gap"
columns that ship in the two input CSVs are OECD's own published numbers for a
*different* AI_i assumption -- they are used **only as a QA cross-check** in
Sections 2-3, never as the source of the numbers we carry forward.

This version is **self-contained**: the AIOE, Complementarity, and C-AIOE
calculation (previously a separate `AIOE_CAIOE_digital_economy.Rmd` run whose
NOC-level output CSV had to be merged in) is now computed inline in Sections
6-8, from the same raw O*NET/mTurk source files that script used. Only one
file needs to be run end to end.

**Required input files (place in the working directory):**

| File | Source | Notes |
|---|---|---|
| `oecd_capability_demands.csv` | OECD AI Exposure Measure public data | ~879 O*NET-SOC occupations; per-domain Job Demands (H_ji) + OECD's own published Capability Gap / Total / Reversed columns (used for QA only) |
| `cognitive_scenario.csv` | Same source, "+1 cognitive capability" scenario export | OECD's own published per-domain gaps under the scenario (QA only -- we recompute the scenario ourselves from `oecd_capability_demands.csv`'s H_ji using scenario AI_i) |
| `abilities.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/abilities.csv | current O*NET Abilities (Importance + Level scales), for AIOE |
| `work_context.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/work_context.csv | for Complementarity |
| `job_zones.csv` | https://www.onetcenter.org/dl_files/database/db_30_3_csv/job_zones.csv | for Complementarity (Skills group) |
| `mturk_ai_relatedness.csv` | AIOE-Data/AIOE GitHub repo (`Input/mturk_mapping_matrix.dta`, exported to CSV) | 10 AI applications x 52 O*NET abilities, crowdsourced relatedness scores, for AIOE |
| `application_list.csv` | Same repo (`Input/application_list.dta`) | flags which of the 10 applications count toward AIOE (all 1 in the base index) |
| `soc_noc_crosswalk.csv` | ICTC crosswalk (SOC -> NOC via The Dais) | `noc, noc_title, onetsoc_code` |

**Note on encoding:** the two OECD CSVs are Windows-1252 encoded (their domain
column headers use an em-dash, `—`), so they are read with an explicit locale
below rather than the default UTF-8 guess. The O*NET/mTurk files are read with
the default UTF-8 guess, matching the original AIOE pipeline.

---

# Section 1 — Load & Harmonize OECD Capability Demand Data

Column headers in the source files are messy (irregular internal whitespace,
em-dashes), so instead of hand-typing exact header strings we match columns by
domain keyword + measure keyword. This is more robust than a positional
rename if OECD ever reorders columns in a refreshed export.

```{r load-demands}
cp1252 <- locale(encoding = "windows-1252")

demands_raw <- read_csv("oecd_capability_demands.csv", locale = cp1252, show_col_types = FALSE)

# Canonical domain order/keys, exactly as OECD AI Papers No. 59 defines them
domain_levels <- c("language", "social_interaction", "problem_solving", "creativity",
                    "metacognition", "knowledge_learning_memory", "vision",
                    "manipulation", "robotic_intelligence")

cognitive_domains <- c("language", "social_interaction", "problem_solving",
                        "creativity", "metacognition", "knowledge_learning_memory")

domain_patterns <- c(
  language                  = "^language",
  social_interaction        = "^social interaction",
  problem_solving           = "^problem solving",
  creativity                = "^creativity",
  metacognition             = "^metacognition",
  knowledge_learning_memory = "^knowledge",
  vision                    = "^vision",
  manipulation              = "^manipulation",
  robotic_intelligence      = "^robotic intelligence"
)

# Classify each raw column name -> (domain, measure) based on squished/lowercased text
classify_domain_column <- function(col_names) {
  cleaned <- col_names %>% str_squish() %>% str_to_lower()
  domain <- rep(NA_character_, length(cleaned))
  for (key in names(domain_patterns)) {
    hit <- str_detect(cleaned, domain_patterns[[key]])
    domain[hit] <- key
  }
  # Only classify a measure if the column also matched a domain -- otherwise
  # summary columns like "AI Capability Gap Index (Total)" (which contains
  # the substring "capability gap" but isn't a per-domain column) get
  # misclassified.
  measure <- case_when(
    is.na(domain) ~ NA_character_,
    str_detect(cleaned, "capability gap") ~ "gap",
    str_detect(cleaned, "job demands")    ~ "demand",
    TRUE ~ NA_character_
  )
  tibble(raw_name = col_names, domain = domain, measure = measure)
}

demands_col_map <- classify_domain_column(names(demands_raw))

# Sanity check: expect exactly 9 domains x 2 measures = 18 domain columns, all classified
n_domain_gap    <- sum(demands_col_map$measure == "gap", na.rm = TRUE)
n_domain_demand <- sum(demands_col_map$measure == "demand", na.rm = TRUE)
cat("Domain gap columns matched:", n_domain_gap, "/ 9\n")
cat("Domain demand columns matched:", n_domain_demand, "/ 9\n")

if (n_domain_gap != 9 || n_domain_demand != 9) {
  warning("Domain column classification did not find exactly 9+9 domain columns -- ",
          "inspect demands_col_map before continuing.")
  print(demands_col_map)
}

# Build a clean, renamed table: onetsoc_code, occupation_title, oecd_reversed_published,
# oecd_total_published, gap_<domain> (OECD published), demand_<domain> (H_ji)
rename_lookup <- demands_col_map %>%
  filter(!is.na(measure)) %>%
  mutate(new_name = paste0(measure, "_", domain)) %>%
  select(new_name, raw_name) %>%   # deframe() uses col 1 as names -> must be the new names
  deframe()

demands <- demands_raw %>%
  rename(
    onetsoc_code            = `Occupation Code (O*NET-SOC Code)`,
    occupation_title        = `Occupation Title`,
    oecd_reversed_published = `AI Capability Gap Index (Reversed, norm.)`,
    oecd_total_published    = `AI Capability Gap Index (Total)`
  ) %>%
  rename(!!!rename_lookup) %>%
  mutate(onetsoc_code = str_trim(onetsoc_code)) %>%
  filter(!is.na(onetsoc_code), onetsoc_code != "")

cat("Occupations loaded:", nrow(demands), "\n")
cat("Duplicate onetsoc_code rows:", sum(duplicated(demands$onetsoc_code)), "\n")

head(demands %>% select(onetsoc_code, occupation_title, starts_with("demand_")))
```

---

# Section 2 — Baseline AI Capability Levels (AI_i) and Domain-Level Capability Gap

Baseline current AI capability levels per domain (0-5 scale), as supplied
directly by the user for this pipeline run. Edit this tribble to update the
assumption without touching any downstream code.

```{r ai-capability-baseline}
ai_capability_baseline <- tribble(
  ~domain,                      ~ai_level_baseline,
  "language",                   3,
  "social_interaction",         2,
  "problem_solving",            2,
  "creativity",                 3,
  "metacognition",               2,
  "knowledge_learning_memory",  2,
  "vision",                     3,
  "manipulation",               3,
  "robotic_intelligence",       2
)

stopifnot(setequal(ai_capability_baseline$domain, domain_levels))
ai_capability_baseline
```

Compute `CG_ji = max(0, H_ji - AI_i)` for every occupation x domain pair, by
pivoting the demand columns to long form, joining the AI_i table, then pivoting
back to one gap column per domain.

```{r domain-gap-baseline}
demand_long <- demands %>%
  select(onetsoc_code, starts_with("demand_")) %>%
  pivot_longer(-onetsoc_code, names_to = "domain", names_prefix = "demand_",
               values_to = "h_ji")

gap_baseline_long <- demand_long %>%
  inner_join(ai_capability_baseline, by = "domain") %>%
  mutate(cg_ji = pmax(0, h_ji - ai_level_baseline))

# Sanity check: join should not drop any domain x occupation rows
if (nrow(gap_baseline_long) != nrow(demand_long)) {
  warning("Baseline AI_i join dropped rows -- check for domain name mismatches. ",
          "Before: ", nrow(demand_long), " After: ", nrow(gap_baseline_long))
}

gap_baseline_wide <- gap_baseline_long %>%
  select(onetsoc_code, domain, cg_ji) %>%
  pivot_wider(names_from = domain, values_from = cg_ji, names_prefix = "gap_calc_")

demands <- demands %>% left_join(gap_baseline_wide, by = "onetsoc_code")

cat("Baseline domain gaps computed for", nrow(gap_baseline_wide), "occupations\n")
```

**QA check** — compare our recomputed `gap_calc_*` against OECD's own published
`gap_*` columns (computed under OECD's own AI_i assumption, which need not match
ours). Large systematic differences are expected if our AI_i differs from
theirs; this check is about catching *join/logic* errors, not enforcing an
exact match.

```{r qa-domain-gap-baseline}
qa_gap_baseline <- purrr::map_dfr(domain_levels, function(d) {
  calc_col <- demands[[paste0("gap_calc_", d)]]
  pub_col  <- demands[[paste0("gap_", d)]]
  tibble(
    domain           = d,
    correlation      = suppressWarnings(cor(calc_col, pub_col, use = "complete.obs")),
    mean_abs_diff    = mean(abs(calc_col - pub_col), na.rm = TRUE),
    exact_match_pct  = mean(abs(calc_col - pub_col) < 1e-6, na.rm = TRUE) * 100
  )
})

print(qa_gap_baseline)
```

---

# Section 3 — Scenario AI Capability Levels (+1 Cognitive) and Domain-Level Gap

The scenario raises AI capability by +1 level (capped at 5) in the 6
**cognitive** domains only (Language, Social Interaction, Problem Solving,
Creativity, Metacognition & Critical Thinking, Knowledge/Learning/Memory).
The 3 physical/robotic domains (Vision, Manipulation, Robotic Intelligence)
are left at their baseline level. Occupational demand (H_ji) does not change
between scenarios -- we reuse the same `demand_*` columns from Section 1.

```{r ai-capability-scenario}
ai_capability_scenario <- ai_capability_baseline %>%
  mutate(
    ai_level_scenario = if_else(domain %in% cognitive_domains,
                                 pmin(5, ai_level_baseline + 1),
                                 ai_level_baseline)
  ) %>%
  select(domain, ai_level_scenario)

ai_capability_scenario
```

```{r domain-gap-scenario}
gap_scenario_long <- demand_long %>%
  inner_join(ai_capability_scenario, by = "domain") %>%
  mutate(cg_ji = pmax(0, h_ji - ai_level_scenario))

if (nrow(gap_scenario_long) != nrow(demand_long)) {
  warning("Scenario AI_i join dropped rows -- check for domain name mismatches. ",
          "Before: ", nrow(demand_long), " After: ", nrow(gap_scenario_long))
}

gap_scenario_wide <- gap_scenario_long %>%
  select(onetsoc_code, domain, cg_ji) %>%
  pivot_wider(names_from = domain, values_from = cg_ji, names_prefix = "gap_scenario_")

demands <- demands %>% left_join(gap_scenario_wide, by = "onetsoc_code")

cat("Scenario domain gaps computed for", nrow(gap_scenario_wide), "occupations\n")

# Sanity check: physical domains must be identical between baseline and scenario,
# since scenario AI_i is unchanged for them
physical_domains <- setdiff(domain_levels, cognitive_domains)
physical_diff <- purrr::map_dbl(physical_domains, function(d) {
  max(abs(demands[[paste0("gap_calc_", d)]] - demands[[paste0("gap_scenario_", d)]]), na.rm = TRUE)
})
names(physical_diff) <- physical_domains
cat("Max |baseline - scenario| gap for physical domains (should be ~0):\n")
print(physical_diff)
```

**QA check** against OECD's own published scenario export
(`cognitive_scenario.csv`), same caveat as Section 2: differences reflect our
different AI_i assumption, not a bug, unless the physical-domain columns
(which should be unaffected by the scenario on both sides) disagree sharply.

```{r qa-domain-gap-scenario}
scenario_raw <- read_csv("cognitive_scenario.csv", locale = cp1252, show_col_types = FALSE)
scenario_col_map <- classify_domain_column(names(scenario_raw))

scenario_rename_lookup <- scenario_col_map %>%
  filter(!is.na(measure), measure == "gap") %>%
  mutate(new_name = paste0("gap_pub_scenario_", domain)) %>%
  select(new_name, raw_name) %>%   # deframe() uses col 1 as names -> must be the new names
  deframe()

scenario_pub <- scenario_raw %>%
  rename(onetsoc_code = `Occupation Code (O*NET-SOC Code)`) %>%
  rename(!!!scenario_rename_lookup) %>%
  mutate(onetsoc_code = str_trim(onetsoc_code)) %>%
  select(onetsoc_code, starts_with("gap_pub_scenario_"))

# Coerce/verify join keys before merging (both should be plain O*NET-SOC strings)
cat("scenario_pub onetsoc_code class:", class(scenario_pub$onetsoc_code), "\n")
cat("demands onetsoc_code class:     ", class(demands$onetsoc_code), "\n")

qa_scenario <- demands %>%
  select(onetsoc_code, starts_with("gap_scenario_")) %>%
  inner_join(scenario_pub, by = "onetsoc_code")

cat("QA scenario match rate:", round(100 * nrow(qa_scenario) / nrow(demands), 1), "%\n")

qa_gap_scenario <- purrr::map_dfr(domain_levels, function(d) {
  calc_col <- qa_scenario[[paste0("gap_scenario_", d)]]
  pub_col  <- qa_scenario[[paste0("gap_pub_scenario_", d)]]
  tibble(
    domain          = d,
    correlation     = suppressWarnings(cor(calc_col, pub_col, use = "complete.obs")),
    mean_abs_diff   = mean(abs(calc_col - pub_col), na.rm = TRUE),
    exact_match_pct = mean(abs(calc_col - pub_col) < 1e-6, na.rm = TRUE) * 100
  )
})

print(qa_gap_scenario)
```

---

# Section 4 — Aggregating Domain Gaps into a Total Capability Gap Index (CG_j)

We don't have OECD's per-occupation importance weights, so we build two
documented approximations and compare them before choosing one:

- **(a) Equal-weighted** — every domain gets weight 1/9 (`Importance_ji` treated
  as uniform across domains, ignoring any occupation-specific variation). This
  is the "no information" fallback, clearly flagged as an approximation.
- **(b) Regression-fit global weights** — OECD's own published Total Index
  (`oecd_total_published`, computed under *their* AI_i, not ours) gives us a
  real target value for all 879 occupations. We fit a single set of
  non-negative global weights `w_i` that best reproduce that target from our
  *own* baseline domain gaps (`gap_calc_*`), via constrained least squares.
  This is far more defensible than eyeballing a handful of published totals,
  even though the weights are still global (not occupation-specific) and the
  regression target used a different AI_i assumption than ours.

```{r weight-fit-equal}
gap_cols_calc <- paste0("gap_calc_", domain_levels)

cg_equal <- demands %>%
  transmute(onetsoc_code, cg_total_equal = rowMeans(across(all_of(gap_cols_calc)), na.rm = TRUE))

demands <- demands %>% left_join(cg_equal, by = "onetsoc_code")

r2_equal <- cor(demands$cg_total_equal, demands$oecd_total_published, use = "complete.obs")^2
cat("Equal-weighted CG_j vs OECD published Total -- R^2 =", round(r2_equal, 3), "\n")
```

```{r weight-fit-regression}
# Non-negative least squares via optim(): minimize SSE of
# oecd_total_published ~ sum_i w_i * gap_calc_i, subject to w_i >= 0
X <- as.matrix(demands[, gap_cols_calc])
y <- demands$oecd_total_published

complete_rows <- complete.cases(X, y)
X_fit <- X[complete_rows, ]
y_fit <- y[complete_rows]

sse <- function(w) sum((y_fit - X_fit %*% w)^2)

fit <- optim(
  par     = rep(1, ncol(X_fit)),
  fn      = sse,
  method  = "L-BFGS-B",
  lower   = rep(0, ncol(X_fit))
)

fitted_weights <- setNames(fit$par, domain_levels)
cat("Fitted global weights (unnormalized, non-negative least squares):\n")
print(round(fitted_weights, 3))

cg_regression <- as.numeric(X %*% fit$par)
demands$cg_total_regression <- cg_regression

r2_regression <- cor(demands$cg_total_regression, demands$oecd_total_published, use = "complete.obs")^2
cat("Regression-fit CG_j vs OECD published Total -- R^2 =", round(r2_regression, 3), "\n")

cat("\nComparison:\n")
cat("  Equal-weighted      R^2 =", round(r2_equal, 3), "\n")
cat("  Regression-fit      R^2 =", round(r2_regression, 3), "\n")
```

**Choice used downstream.** Set `weighting_method` below to `"regression"` or
`"equal"` to control which total index is carried forward through the rest of
the pipeline. Both remain in the output table for comparison regardless.

```{r weight-fit-choice}
weighting_method <- "regression"   # or "equal"

demands <- demands %>%
  mutate(
    cg_total_baseline = if (weighting_method == "regression") cg_total_regression else cg_total_equal
  )

# Apply the same chosen weights (equal or regression-fit) to the scenario gaps
gap_cols_scenario <- paste0("gap_scenario_", domain_levels)

if (weighting_method == "regression") {
  demands$cg_total_scenario <- as.numeric(as.matrix(demands[, gap_cols_scenario]) %*% fit$par)
} else {
  demands$cg_total_scenario <- rowMeans(demands[, gap_cols_scenario], na.rm = TRUE)
}

cat("Using weighting_method =", weighting_method, "\n")
cat("Baseline CG_j range: min =", round(min(demands$cg_total_baseline, na.rm = TRUE), 2),
    " max =", round(max(demands$cg_total_baseline, na.rm = TRUE), 2), "\n")
cat("Scenario CG_j range: min =", round(min(demands$cg_total_scenario, na.rm = TRUE), 2),
    " max =", round(max(demands$cg_total_scenario, na.rm = TRUE), 2), "\n")
```

---

# Section 5 — Reversed / Normalized Exposure Score

`exposure = 1 - CG_j / 23`, matching OECD's convention (23 is their documented
effective real-world maximum of the total index). Higher = smaller capability
gap = AI more capable relative to the job's requirements = **more exposed**.

```{r reversed-score}
gap_index_max <- 23

demands <- demands %>%
  mutate(
    exposure_reversed_baseline = pmax(0, 1 - cg_total_baseline / gap_index_max),
    exposure_reversed_scenario = pmax(0, 1 - cg_total_scenario / gap_index_max)
  )

# Sanity check: flag any occupation whose CG_j exceeds the assumed 23 max
n_over_max <- sum(demands$cg_total_baseline > gap_index_max, na.rm = TRUE)
if (n_over_max > 0) {
  warning(n_over_max, " occupations have CG_j > ", gap_index_max,
          " -- the 23 cap may not hold for this weighting; reversed score was clamped at 0.")
}

cat("Baseline exposure_reversed range: min =", round(min(demands$exposure_reversed_baseline), 3),
    " max =", round(max(demands$exposure_reversed_baseline), 3), "\n")

write_csv(demands, "oecd_capability_gap_occupation_level.csv")
```

---

# Section 6 — AIOE: Ability-Level Exposure (A_j) and Occupation-Level AIOE

Ported from `AIOE_CAIOE_digital_economy.Rmd` Sections 1-2, unchanged. This
reproduces Data Appendix E of Felten, Raj & Seamans (2021): for each of the 52
O*NET abilities, sum its crowdsourced relatedness to the 10 AI applications,
then standardize (z-score) across abilities. This step uses **only** the
crowdsourced mTurk matrix -- it does not depend on any O*NET occupation data.

```{r ability-exposure}
mturk <- read_csv("mturk_ai_relatedness.csv", show_col_types = FALSE)
app_list <- read_csv("application_list.csv", show_col_types = FALSE)

# Ability columns are everything except application_id / applications
ability_cols <- setdiff(names(mturk), c("application_id", "applications"))

# Multiply each application's relatedness row by its inclusion flag (binary_all),
# then sum across all 10 applications for each ability -> raw ability exposure
mturk_weighted <- mturk %>%
  left_join(app_list %>% select(applications, score = binary_all), by = "applications") %>%
  mutate(across(all_of(ability_cols), ~ .x * score))

ability_exposure_raw <- mturk_weighted %>%
  summarise(across(all_of(ability_cols), sum)) %>%
  pivot_longer(everything(), names_to = "onet_element", values_to = "onet_score")

# Standardize (z-score) across the 52 abilities -- this is "ability_exposure"
ability_exposure <- ability_exposure_raw %>%
  mutate(ability_exposure = as.numeric(scale(onet_score))) %>%
  select(onet_element, ability_exposure)

head(ability_exposure)
```

```{r aioe-occupation}
abilities <- read_csv("abilities.csv", show_col_types = FALSE)

abilities_clean <- abilities %>%
  rename(onetsoc_code = `O*NET-SOC Code`,
         element_id   = `Element ID`,
         element_name = `Element Name`,
         scale_id     = `Scale ID`,
         data_value   = `Data Value`) %>%
  mutate(
    onet_element = str_remove_all(str_to_lower(element_name), "[ -]")
  )

# Rescale Importance (IM, native 1-5) and Level (LV, native 1-7), per the
# original methodology, then compute scalar = level_scaled x importance_scaled
scaled <- abilities_clean %>%
  filter(scale_id %in% c("IM", "LV")) %>%
  mutate(
    scaled_value = case_when(
      scale_id == "IM" ~ data_value / 5,
      scale_id == "LV" ~ data_value / 7
    )
  ) %>%
  select(onetsoc_code, element_id, onet_element, scale_id, scaled_value) %>%
  pivot_wider(names_from = scale_id, values_from = scaled_value, values_fn = max) %>%
  rename(importance_scaled = IM, level_scaled = LV) %>%
  filter(!is.na(importance_scaled), !is.na(level_scaled)) %>%
  mutate(scalar = level_scaled * importance_scaled)

# Sanity check: every mTurk ability should exist in the current O*NET release
missing_abilities <- setdiff(ability_exposure$onet_element, unique(scaled$onet_element))
if (length(missing_abilities) > 0) {
  warning("Abilities present in mTurk matrix but missing from current O*NET data: ",
          paste(missing_abilities, collapse = "; "))
}

# Join ability exposure onto the occupation-ability table, weight, and sum
occ_ability <- scaled %>%
  inner_join(ability_exposure, by = "onet_element") %>%
  mutate(wtd_element_impact = scalar * ability_exposure)

occ_aioe_detailed <- occ_ability %>%
  group_by(onetsoc_code) %>%
  summarise(
    ability_base          = sum(scalar, na.rm = TRUE),
    total_weighted_impact = sum(wtd_element_impact, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(ability_base_impact = total_weighted_impact / ability_base)

# Aggregate detailed O*NET-SOC (8-digit) up to base SOC (7-char, e.g. 15-1252),
# averaging across detailed occupations that share a base SOC code
occ_aioe_soc <- occ_aioe_detailed %>%
  mutate(onetsoc_base = substr(onetsoc_code, 1, 7)) %>%
  group_by(onetsoc_base) %>%
  summarise(ability_base_impact = mean(ability_base_impact, na.rm = TRUE), .groups = "drop") %>%
  mutate(aioe_score = as.numeric(scale(ability_base_impact))) %>%   # final standardization
  arrange(desc(aioe_score))

cat("AIOE computed for", nrow(occ_aioe_soc), "SOC occupations\n")
cat("Range: min =", round(min(occ_aioe_soc$aioe_score), 3),
    " max =", round(max(occ_aioe_soc$aioe_score), 3), "\n")
```

---

# Section 7 — Complementarity Score (O*NET Work Context)

Ported from `AIOE_CAIOE_digital_economy.Rmd` Section 3, unchanged.

```{r complementarity}
work_context <- read_csv("work_context.csv", show_col_types = FALSE)
job_zones    <- read_csv("job_zones.csv",    show_col_types = FALSE)

cx <- work_context %>%
  filter(`Scale ID` == "CX") %>%
  select(onetsoc_code = `O*NET-SOC Code`,
         element_name = `Element Name`,
         data_value   = `Data Value`)

element_scales <- tribble(
  ~element_name,                                                  ~scale_min, ~scale_max, ~group,
  "Face-to-Face Discussions with Individuals and Within Teams",   1,          5,          "communication",
  "Public Speaking",                                              1,          5,          "communication",
  "Work Outcomes and Results of Other Workers",                   1,          5,          "responsibility",
  "Health and Safety of Other Workers",                           1,          5,          "responsibility",
  "Outdoors, Exposed to All Weather Conditions",                  1,          5,          "physical",
  "Physical Proximity",                                           1,          5,          "physical",
  "Consequence of Error",                                         1,          5,          "criticality",
  "Freedom to Make Decisions",                                    1,          5,          "criticality",
  "Frequency of Decision Making",                                 1,          5,          "criticality",
  "Degree of Automation",                                         0,          100,        "routine",
  "Importance of Repeating Same Tasks",                           1,          5,          "routine",
)

missing_elements <- setdiff(element_scales$element_name, unique(cx$element_name))
if (length(missing_elements) > 0) {
  warning("These element names were not found in Work Context.csv: ",
          paste(missing_elements, collapse = "; "))
}

cx_scaled <- cx %>%
  inner_join(element_scales, by = "element_name") %>%
  mutate(value_0_100 = (data_value - scale_min) / (scale_max - scale_min) * 100) %>%
  mutate(value_0_100 = if_else(element_name == "Degree of Automation",
                                100 - value_0_100, value_0_100))

group_means <- cx_scaled %>%
  group_by(onetsoc_code, group) %>%
  summarise(group_mean = mean(value_0_100, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = group, values_from = group_mean, names_glue = "{group}_mean")

job_zones_scaled <- job_zones %>%
  select(onetsoc_code = `O*NET-SOC Code`, job_zone = `Job Zone`) %>%
  mutate(skills_mean = job_zone * 20)

complementarity <- group_means %>%
  left_join(job_zones_scaled %>% select(onetsoc_code, skills_mean), by = "onetsoc_code") %>%
  rowwise() %>%
  mutate(
    complementarity_score = mean(c(communication_mean, responsibility_mean,
                                    physical_mean, criticality_mean,
                                    routine_mean, skills_mean), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}")) %>%
  arrange(desc(complementarity_score))

cat("Complementarity computed for", nrow(complementarity), "occupations\n")
cat("Range: min =", round(min(complementarity$complementarity_score, na.rm = TRUE), 1),
    " max =", round(max(complementarity$complementarity_score, na.rm = TRUE), 1), "\n")
```

---

# Section 8 — Merge AIOE + Complementarity, Compute C-AIOE

Ported from `AIOE_CAIOE_digital_economy.Rmd` Section 4, unchanged. Following
Pizzinelli et al. (2023, IMF WP/23/216), equation (1), using their literal
convention (theta already on a 0-1 basis, so divide by 100 rather than by the
empirical range):

`C-AIOE_i = AIOE_i * (1 - (theta_i - theta_MIN) / 100)`

```{r merge-caioe}
aioe_comp <- occ_aioe_soc %>%
  select(onetsoc_base, aioe_score) %>%
  distinct() %>%
  inner_join(
    complementarity %>% select(onetsoc_base, complementarity_score) %>% distinct(),
    by = "onetsoc_base"
  )

theta_min <- min(aioe_comp$complementarity_score, na.rm = TRUE)

aioe_comp <- aioe_comp %>%
  mutate(
    c_aioe = aioe_score * (1 - (complementarity_score - theta_min) / 100)
  )

# ---- Augmentation Index (range-anchored, mirrors C-AIOE's theta_MIN anchoring) ----
comp_min <- min(aioe_comp$complementarity_score, na.rm = TRUE)
comp_max <- max(aioe_comp$complementarity_score, na.rm = TRUE)

aioe_comp <- aioe_comp %>%
  mutate(
    augmentation_index = aioe_score * ((complementarity_score - comp_min) / (comp_max - comp_min))
  )

cat("theta_MIN =", round(theta_min, 2), "\n")
head(aioe_comp)
```

---

# Section 9 — SOC -> NOC Aggregation: OECD Gap Index

Same three-tier pattern used throughout: (1) direct match on 6-digit base SOC,
(2) backward-crosswalk fallback for any SOC code revisions, (3) major (2-digit)
SOC-group average as a last resort for anything still unmatched. Coerce join
keys and verify row counts at each step. `crosswalk` is loaded once here and
reused in Section 10 for the AIOE/Complementarity aggregation.

```{r crosswalk-load}
crosswalk <- read_csv("soc_noc_crosswalk.csv", show_col_types = FALSE,
                       col_types = cols(noc = col_character())) %>%
  mutate(
    onetsoc_code = str_trim(as.character(onetsoc_code)),
    onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}")
  )

demands <- demands %>%
  mutate(onetsoc_base = str_extract(onetsoc_code, "^[0-9]{2}-[0-9]{4}"))

cat("Crosswalk rows:", nrow(crosswalk), " | Distinct NOCs:", n_distinct(crosswalk$noc), "\n")
cat("Distinct onetsoc_base in gap data:", n_distinct(demands$onetsoc_base), "\n")
```

```{r crosswalk-tier1-direct}
# ---- Tier 1: direct match on base SOC code ---------------------------------
# relationship = "many-to-many" is expected and intentional here: a base SOC
# code can cover several detailed O*NET-SOC occupations in `demands` (e.g.
# 11-1011.00 and 11-1011.03 both -> base 11-1011), and each should contribute
# to the NOC-level average computed later.
matched_direct <- crosswalk %>%
  inner_join(
    demands %>% select(onetsoc_base, cg_total_baseline, cg_total_scenario,
                        exposure_reversed_baseline, exposure_reversed_scenario),
    by = "onetsoc_base", relationship = "many-to-many"
  ) %>%
  mutate(match_method = "direct")

unmatched_after_tier1 <- crosswalk %>%
  anti_join(demands, by = "onetsoc_base")

cat("Tier 1 (direct): ", n_distinct(matched_direct$onetsoc_code), " of ",
    n_distinct(crosswalk$onetsoc_code), " O*NET-SOC codes matched\n")
cat("Unmatched after tier 1:", n_distinct(unmatched_after_tier1$onetsoc_code), "\n")
```

```{r crosswalk-tier2-backward}
# ---- Tier 2: backward crosswalk for revised/retired SOC codes -------------
# Populate this tribble with known SOC-vintage pairs (old_onetsoc_base ->
# current_onetsoc_base) if/when unmatched codes turn out to be revisions
# rather than genuine gaps. Empty by default -- extend as needed.
# Explicit character(0) typing avoids an empty tribble defaulting to logical
# columns, which would otherwise fail the join below with a type mismatch.
soc_backward_crosswalk <- tibble(
  old_onetsoc_base     = character(0),
  current_onetsoc_base = character(0)
  # e.g. add rows: tibble(old_onetsoc_base = "15-1141", current_onetsoc_base = "15-1152")
)

matched_tier2 <- unmatched_after_tier1 %>%
  inner_join(soc_backward_crosswalk, by = c("onetsoc_base" = "old_onetsoc_base")) %>%
  inner_join(
    demands %>% select(onetsoc_base, cg_total_baseline, cg_total_scenario,
                        exposure_reversed_baseline, exposure_reversed_scenario),
    by = c("current_onetsoc_base" = "onetsoc_base"), relationship = "many-to-many"
  ) %>%
  mutate(match_method = "backward_crosswalk") %>%
  select(-current_onetsoc_base)

unmatched_after_tier2 <- unmatched_after_tier1 %>%
  anti_join(soc_backward_crosswalk, by = c("onetsoc_base" = "old_onetsoc_base"))

cat("Tier 2 (backward crosswalk): ", n_distinct(matched_tier2$onetsoc_code), " additional codes matched\n")
cat("Unmatched after tier 2:", n_distinct(unmatched_after_tier2$onetsoc_code), "\n")
```

```{r crosswalk-tier3-majorgroup}
# ---- Tier 3: 2-digit major SOC group average, last resort -----------------
major_group_avg <- demands %>%
  mutate(major_group = str_extract(onetsoc_base, "^[0-9]{2}")) %>%
  group_by(major_group) %>%
  summarise(
    cg_total_baseline           = mean(cg_total_baseline, na.rm = TRUE),
    cg_total_scenario           = mean(cg_total_scenario, na.rm = TRUE),
    exposure_reversed_baseline  = mean(exposure_reversed_baseline, na.rm = TRUE),
    exposure_reversed_scenario  = mean(exposure_reversed_scenario, na.rm = TRUE),
    .groups = "drop"
  )

matched_tier3 <- unmatched_after_tier2 %>%
  mutate(major_group = str_extract(onetsoc_base, "^[0-9]{2}")) %>%
  inner_join(major_group_avg, by = "major_group") %>%
  mutate(match_method = "major_group_average") %>%
  select(-major_group)

still_unmatched <- unmatched_after_tier2 %>%
  mutate(major_group = str_extract(onetsoc_base, "^[0-9]{2}")) %>%
  anti_join(major_group_avg, by = "major_group")

cat("Tier 3 (major group average): ", n_distinct(matched_tier3$onetsoc_code), " additional codes matched\n")
cat("Still unmatched (no major group fallback available):", n_distinct(still_unmatched$onetsoc_code), "\n")
```

```{r crosswalk-noc-aggregate}
# ---- Combine all tiers and aggregate to NOC --------------------------------
all_matched <- bind_rows(matched_direct, matched_tier2, matched_tier3)

# Sanity check: a given (onetsoc_code, noc) pair may legitimately appear more
# than once (base-SOC fan-out to multiple detailed occupations, see note
# above) -- what should never happen is the *same* pair being matched via
# more than one tier (e.g. both "direct" and "major_group_average"), which
# would mean the tiering logic isn't mutually exclusive.
dup_check <- all_matched %>%
  group_by(onetsoc_code, noc) %>%
  summarise(n_methods = n_distinct(match_method), .groups = "drop") %>%
  filter(n_methods > 1)
if (nrow(dup_check) > 0) {
  warning(nrow(dup_check), " onetsoc_code x noc combinations matched via more than one tier -- check for overlap.")
}

# Note: gap data here is not pre-collapsed to base SOC before this join, so
# where a base SOC covers multiple detailed O*NET-SOC occupations (e.g.
# 11-1011.00 and 11-1011.03 both -> base 11-1011), the join fans out and
# n_onet_linked_oecd below counts each matched detailed occupation -- this is
# intentional, so the NOC-level mean is an average across every linked
# detailed occupation, not just distinct base codes.
noc_oecd_gap <- all_matched %>%
  group_by(noc, noc_title) %>%
  summarise(
    oecd_cg_total_baseline          = mean(cg_total_baseline, na.rm = TRUE),
    oecd_cg_total_scenario          = mean(cg_total_scenario, na.rm = TRUE),
    oecd_exposure_reversed_baseline = mean(exposure_reversed_baseline, na.rm = TRUE),
    oecd_exposure_reversed_scenario = mean(exposure_reversed_scenario, na.rm = TRUE),
    n_onet_linked_oecd              = n(),
    n_direct                        = sum(match_method == "direct"),
    n_backward_crosswalk            = sum(match_method == "backward_crosswalk"),
    n_major_group_fallback          = sum(match_method == "major_group_average"),
    .groups = "drop"
  ) %>%
  arrange(desc(oecd_exposure_reversed_baseline))

cat("NOC-level OECD gap results:", nrow(noc_oecd_gap), "occupations\n")
cat("NOCs relying on backward crosswalk:", sum(noc_oecd_gap$n_backward_crosswalk > 0), "\n")
cat("NOCs relying on major-group fallback:", sum(noc_oecd_gap$n_major_group_fallback > 0), "\n")
cat("Coverage: ", round(100 * nrow(noc_oecd_gap) / n_distinct(crosswalk$noc), 1),
    "% of crosswalk NOCs have at least one linked O*NET-SOC code\n")

write_csv(noc_oecd_gap, "noc_oecd_capability_gap.csv")
```

---

# Section 10 — SOC -> NOC Aggregation: AIOE / Complementarity / C-AIOE

Ported from `AIOE_CAIOE_digital_economy.Rmd` Sections 5-6 (crosswalk join,
median-split quadrant, digital economy tier lookup), unchanged, but reusing
the same `crosswalk` table loaded in Section 9 instead of re-reading the file.
This uses a plain direct join (not the three-tier fallback above), matching
the original AIOE pipeline's own convention.

```{r crosswalk-aioe}
noc_aioe_comp <- crosswalk %>%
  inner_join(aioe_comp, by = "onetsoc_base") %>%
  group_by(noc, noc_title) %>%
  summarise(
    aioe_score             = mean(aioe_score, na.rm = TRUE),
    complementarity_score  = mean(complementarity_score, na.rm = TRUE),
    c_aioe                 = mean(c_aioe, na.rm = TRUE),
    augmentation_index     = mean(augmentation_index, na.rm = TRUE),
    n_onet_linked          = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(c_aioe))

cat("NOC-level AIOE/Complementarity results:", nrow(noc_aioe_comp), "occupations\n")
```

```{r aioe-quadrants}
aioe_median <- median(noc_aioe_comp$aioe_score, na.rm = TRUE)
comp_median <- median(noc_aioe_comp$complementarity_score, na.rm = TRUE)

noc_aioe_comp <- noc_aioe_comp %>%
  mutate(
    quadrant = case_when(
      aioe_score >= aioe_median & complementarity_score >= comp_median ~
        "High Exposure, High Complementarity",
      aioe_score >= aioe_median & complementarity_score < comp_median ~
        "High Exposure, Low Complementarity",
      TRUE ~ "Low Exposure"
    )
  )

cat("AIOE median:", round(aioe_median, 3), " | Complementarity median:", round(comp_median, 2), "\n")
table(noc_aioe_comp$quadrant)

# ---- Digital Economy Tier Lookup ----------------------------------------
digital_economy_tiers <- tribble(
  ~noc,    ~noc_desc,                                                      ~tier,
  "00014", "Senior managers - trade, broadcasting",                       "Digital-Enabled",
  "10022", "Advertising, marketing, PR managers",                         "Digital-Enabled",
  "10030", "Telecommunication carriers managers",                         "Core-Digital",
  "12111", "Health information management",                               "Digital-Enabled",
  "12112", "Records management technicians",                              "Digital-Enabled",
  "14112", "Desktop publishing operators",                                "Digital-Enabled",
  "20010", "Engineering managers",                                        "Digital-Intensive",
  "20012", "Computer and information systems managers",                   "Core-Digital",
  "11202", "Professional occupations in advertising, marketing and PR",   "Digital-Enabled",
  "21100", "Physicists and astronomers",                                  "Digital-Intensive",
  "21200", "Architects",                                                  "Digital-Intensive",
  "21210", "Mathematicians, statisticians, actuaries",                    "Digital-Enabled",
  "21211", "Data scientists",                                             "Core-Digital",
  "21220", "Cybersecurity specialists",                                   "Core-Digital",
  "21221", "Business systems specialists",                                "Core-Digital",
  "21222", "Information systems specialists",                             "Core-Digital",
  "21223", "Database analysts and data administrators",                   "Core-Digital",
  "21230", "Computer systems developers",                                 "Core-Digital",
  "21231", "Software engineers and designers",                            "Core-Digital",
  "21232", "Software developers and programmers",                        "Core-Digital",
  "21233", "Web designers",                                               "Core-Digital",
  "21234", "Web developers and programmers",                              "Core-Digital",
  "21300", "Civil engineers",                                             "Digital-Intensive",
  "21301", "Mechanical engineers",                                        "Digital-Intensive",
  "21310", "Electrical and electronics engineers",                        "Core-Digital",
  "21311", "Computer engineers",                                          "Core-Digital",
  "21321", "Industrial/manufacturing engineers",                          "Digital-Intensive",
  "21390", "Aerospace engineers",                                         "Digital-Intensive",
  "22210", "Architectural technologists",                                 "Digital-Intensive",
  "22211", "Industrial designers",                                        "Digital-Intensive",
  "22212", "Drafting technologists and technicians",                      "Digital-Enabled",
  "22214", "Geomatics and meteorology technicians",                       "Digital-Enabled",
  "22220", "Computer network and web technicians",                        "Core-Digital",
  "22221", "User support technicians",                                    "Core-Digital",
  "22222", "Information systems testing technicians",                     "Core-Digital",
  "22301", "Mechanical engineering technologists",                        "Digital-Intensive",
  "22302", "Industrial engineering technologists",                        "Digital-Adjacent",
  "22310", "Electrical/electronics engineering techs",                    "Digital-Adjacent",
  "22311", "Electronic service technicians",                              "Digital-Adjacent",
  "50011", "Managers - publishing, broadcasting",                         "Digital-Enabled",
  "52110", "Film and video camera operators",                             "Digital-Adjacent",
  "52111", "Graphic arts technicians",                                    "Digital-Enabled",
  "52112", "Broadcast technicians",                                       "Digital-Adjacent",
  "52113", "Audio and video recording technicians",                       "Digital-Adjacent",
  "52120", "Graphic designers and illustrators",                          "Digital-Enabled",
  "53111", "Broadcasting, photography assistants",                        "Digital-Adjacent",
  "62100", "Technical sales specialists",                                 "Digital-Adjacent",
  "72011", "Electrical trades supervisors",                               "Digital-Adjacent",
  "72201", "Industrial electricians",                                     "Digital-Adjacent",
  "72202", "Power system electricians",                                   "Digital-Adjacent",
  "72203", "Electrical power line workers",                                "Digital-Adjacent",
  "72204", "Telecommunications line installers",                          "Digital-Adjacent",
  "72205", "Telecommunications equipment installers",                     "Digital-Adjacent",
  "92021", "Electronics manufacturing supervisors",                       "Digital-Adjacent",
  "94201", "Electronics assemblers",                                      "Digital-Adjacent"
) %>%
  select(noc, tier)   # drop noc_desc here, since noc_aioe_comp already has noc_title

# ---- Join tier + derive binary digital economy flag ----------------------
noc_aioe_comp <- noc_aioe_comp %>%
  mutate(noc = str_pad(as.character(noc), width = 5, side = "left", pad = "0"))

digital_economy_tiers <- digital_economy_tiers %>%
  mutate(noc = str_pad(as.character(noc), width = 5, side = "left", pad = "0"))

noc_aioe_comp <- noc_aioe_comp %>%
  left_join(digital_economy_tiers, by = "noc") %>%
  mutate(
    digital_economy = if_else(!is.na(tier), "Digital Economy", "Non-Digital Economy"),
    tier = if_else(is.na(tier), "Not Applicable", tier)
  )

cat("Digital economy classification:\n")
print(table(noc_aioe_comp$digital_economy))
cat("\nTier breakdown:\n")
print(table(noc_aioe_comp$tier))
```

---

# Section 11 — Merge OECD Gap Index with AIOE / Complementarity

The OECD Gap Index is merged in as **parallel columns**, not blended into a
single number, so AIOE / Complementarity / OECD Gap can be compared side by
side per NOC. Both NOC-level tables were computed inline above (Sections 9-10),
so this is a local merge, not a file read.

```{r merge-aioe-complementarity}
noc_oecd_gap <- noc_oecd_gap %>%
  mutate(noc = str_pad(as.character(noc), width = 5, side = "left", pad = "0"))

cat("AIOE/Complementarity NOCs:", n_distinct(noc_aioe_comp$noc), "\n")
cat("OECD gap NOCs:            ", n_distinct(noc_oecd_gap$noc), "\n")

comparison <- noc_aioe_comp %>%
  left_join(
    noc_oecd_gap %>% select(-noc_title),
    by = "noc"
  )

n_matched <- sum(!is.na(comparison$oecd_exposure_reversed_baseline))
cat("Matched to OECD gap data:", n_matched, "of", nrow(comparison), "NOCs (",
    round(100 * n_matched / nrow(comparison), 1), "%)\n")
```

---

# Section 12 — Divergence Flag: OECD Gap Index vs AIOE

Both `aioe_score` and `exposure_reversed_baseline` should tell a broadly
similar story (higher = more exposed), but they come from very different
methodologies. We rank each NOC by percentile within each measure and flag
sharp disagreements -- e.g. AIOE says highly exposed while OECD says lowly
exposed, or vice versa.

```{r divergence-flag}
divergence_threshold <- 0.4   # flag when percentile ranks differ by more than this

comparison <- comparison %>%
  mutate(
    aioe_percentile = percent_rank(aioe_score),
    oecd_percentile  = percent_rank(oecd_exposure_reversed_baseline),
    percentile_gap   = aioe_percentile - oecd_percentile,
    divergence_flag  = abs(percentile_gap) > divergence_threshold & !is.na(percentile_gap),
    divergence_direction = case_when(
      is.na(percentile_gap)  ~ NA_character_,   # no OECD match -- not comparable, not "consistent"
      !divergence_flag       ~ "Consistent",
      percentile_gap > 0     ~ "AIOE High / OECD Low",
      percentile_gap < 0     ~ "AIOE Low / OECD High",
      TRUE ~ NA_character_
    )
  )

cat("NOCs flagged as divergent (|percentile gap| >", divergence_threshold, "):",
    sum(comparison$divergence_flag, na.rm = TRUE), "of",
    sum(!is.na(comparison$percentile_gap)), "comparable NOCs\n")

print(table(comparison$divergence_direction, useNA = "ifany"))

comparison %>%
  filter(divergence_flag) %>%
  select(noc, noc_title, aioe_score, oecd_exposure_reversed_baseline,
         aioe_percentile, oecd_percentile, divergence_direction) %>%
  arrange(desc(abs(aioe_percentile - oecd_percentile))) %>%
  head(15)
```

---

# Section 13 — OECD Quadrant Classification (Median Split)

Same median-split convention as `quadrant` (Section 10 above), applied here to
the OECD exposure score instead of AIOE, so the two quadrant classifications
can be compared side by side per NOC.

```{r oecd-quadrants}
oecd_exposure_median <- median(comparison$oecd_exposure_reversed_baseline, na.rm = TRUE)
comp_median_oecd      <- median(comparison$complementarity_score, na.rm = TRUE)

comparison <- comparison %>%
  mutate(
    oecd_quadrant = case_when(
      is.na(oecd_exposure_reversed_baseline) | is.na(complementarity_score) ~ NA_character_,
      oecd_exposure_reversed_baseline >= oecd_exposure_median & complementarity_score >= comp_median_oecd ~
        "High Exposure, High Complementarity",
      oecd_exposure_reversed_baseline >= oecd_exposure_median & complementarity_score < comp_median_oecd ~
        "High Exposure, Low Complementarity",
      TRUE ~ "Low Exposure"
    )
  )

cat("OECD exposure median:", round(oecd_exposure_median, 3),
    " | Complementarity median:", round(comp_median_oecd, 2), "\n")
table(comparison$oecd_quadrant, useNA = "ifany")
```

---

# Section 14 — Final Output

```{r write-final}
write_csv(comparison, "noc_oecd_gap_vs_aioe_complementarity.csv")

cat("Final comparison table:", nrow(comparison), "NOCs, ", ncol(comparison), "columns\n")
```

---

# Summary

Final output (`noc_oecd_gap_vs_aioe_complementarity.csv`) contains, per NOC:

- `oecd_cg_total_baseline` / `oecd_cg_total_scenario` -- OECD-style total
  Capability Gap Index (`weighting_method` in Section 4 controls whether this
  is equal-weighted or regression-fit)
- `oecd_exposure_reversed_baseline` / `oecd_exposure_reversed_scenario` --
  normalized exposure score, `1 - CG_j / 23`
- `aioe_score`, `complementarity_score`, `c_aioe`, `augmentation_index`,
  `quadrant`, `tier`, `digital_economy` -- computed inline in Sections 6-10
  (previously required a separate `AIOE_CAIOE_digital_economy.Rmd` run and
  file merge)
- `n_onet_linked_oecd`, `n_direct`, `n_backward_crosswalk`,
  `n_major_group_fallback` -- coverage diagnostics for the OECD SOC->NOC mapping
- `n_onet_linked` -- coverage diagnostic for the AIOE/Complementarity SOC->NOC mapping
- `divergence_flag`, `divergence_direction` -- flags NOCs where the OECD
  exposure ranking and the AIOE ranking disagree sharply
- `oecd_quadrant` -- median-split classification of the OECD exposure score
  (High Exposure/High Complementarity, High Exposure/Low Complementarity,
  Low Exposure), same convention as `quadrant`, so the two can be compared
  side by side

Intermediate outputs, useful for auditing:

- `oecd_capability_gap_occupation_level.csv` -- full occupation-level detail
  (domain demands, computed gaps, both weighting schemes, QA columns)
- `noc_oecd_capability_gap.csv` -- OECD gap index at NOC level, before merging
  with AIOE/Complementarity
